---
title: "Networks"
author: "Anthony Rentsch"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    collapsed: false
    smooth_scroll: false
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Dependencies
library(tidyverse)
library(network)
library(sna)
library(igraph)
library(GGally)
library(tictoc)

# Data
orders <- read.csv("/Users/anthonyrentsch/Desktop/UMass/Senior Year/First Semester/Advanced Data Analysis/orders.csv")
order_products_prior <- read.csv("/Users/anthonyrentsch/Desktop/UMass/Senior Year/First Semester/Advanced Data Analysis/order_products__prior.csv")
order_products_train <- read.csv("order_products__train.csv")
products <- read.csv("products.csv")
aisles <- read.csv("aisles.csv")
departments <- read.csv("departments.csv")
```

# Setup

```{r}
#Here, we join the orders by time with the products ordered for prior and train set orders
prior <- order_products_prior %>% left_join(orders, by = "order_id")
train <- order_products_train %>% left_join(orders, by = "order_id")
#Joining the prior and training sets together
orders_full <- rbind(train, prior)
```

# Construct Network

Each product is a node. Each connection runs between products that are in the same order. 

```{r}
# this function calculates all possible pairs from a column of product_ids
# create_edgelist <- function(x) {
#   l <- length(x)
#   mylist <- lapply(2:l, function(y) {
#     combn(x, y, simplify = FALSE)
#   })
#   to.return <- do.call(rbind.data.frame, mylist[[1]])
#   names(to.return) <- c("product1","product2")
#   to.return
# }

# create edgelist matrix for just 3 orders
edgelist <- data.frame()

set.seed(11)
orders_full_small <- orders_full %>% 
  filter(order_id %in% sample(orders_full$order_id, size = 5, replace = T)) %>% 
  select(order_id, product_id)
orders_full_small$order_id <- as.factor(orders_full_small$order_id)

# for(x in levels(orders_full_small$order_id)){
#   temp <- create_edgelist(orders_full_small$product_id[orders_full_small$order_id == x])
#   edgelist <- rbind(edgelist, temp)
# }
# 
# # create igraph network
# tic()
# network_igraph <- graph_from_data_frame(edgelist, directed=FALSE)
# toc()
```

# Visualize network

igraph 

```{r}
# plot(network_igraph)
```

# Summary stats

Look at number of  adjacent verticies that each vertex has. This means we will look at how many different products each product was in an order with - this gives a sense of how widely popular an item is.

```{r}
# this calculates the adjacent edges for every vertex
# degree(network_igraph, V(network_igraph)) %>% 
#   as.data.frame() %>% 
#   rownames_to_column() %>%
#   rename('product_id' = 'rowname', 'num_adjacent_edges' = '.') %>% 
#   mutate(product_id = as.integer(product_id)) %>% 
#   left_join(products, by = "product_id") %>% 
#   select(product_id, product_name, num_adjacent_edges) %>% 
#   arrange(desc(num_adjacent_edges))
```


# Setting up edgelist matrix as a n x n matrix (n = 62)

```{r}
all_product_ids <- unique(orders_full_small$product_id)
edgelist_matrix <- matrix(nrow = length(all_product_ids),
                          ncol = length(all_product_ids),
                          dimnames = list(all_product_ids, all_product_ids))

for (x in levels(orders_full_small$order_id)){
  for (i in rownames(edgelist_matrix)){
    for (j in colnames(edgelist_matrix)){
       if (i %in% unique(orders_full_small$product_id[orders_full_small$order_id == x])){
         if (j %in% unique(orders_full_small$product_id[orders_full_small$order_id == x])){
           edgelist_matrix[i,j] <- 1
         }
         else edgelist_matrix[i,j] <- 0
       }
    }
  }
}

# rename rows
rnames <- rownames(edgelist_matrix) %>% 
  as.data.frame() %>% 
  rename(product_id = '.') %>% 
  mutate(product_id = as.integer(product_id)) %>% 
  left_join(products, by = "product_id") %>% 
  select(product_name)
rnames <- rnames$product_name
rnames <- as.character(rnames)

rownames(edgelist_matrix) <- rnames

# rename columns
cnames <- colnames(edgelist_matrix) %>% 
  as.data.frame() %>% 
  rename(product_id = '.') %>% 
  mutate(product_id = as.integer(product_id)) %>% 
  left_join(products, by = "product_id") %>% 
  select(product_name)
cnames <- cnames$product_name
cnames <- as.character(cnames)

colnames(edgelist_matrix) <- cnames
```

What products do we have here?

```{r}
rownames(edgelist_matrix) %>% 
  as.data.frame() %>% 
  rename(product_id = '.') %>% 
  mutate(product_id = as.integer(product_id)) %>% 
  left_join(products, by = "product_id") %>% 
  select(product_name)
```

Graph from adjacency matrix.

```{r}
igraph_adj_network <- graph_from_adjacency_matrix(edgelist_matrix, mode = "undirected", weighted = TRUE,
                                                  add.colnames = 'label')

# function borrowed from Stack Overflow
# https://stackoverflow.com/questions/38999656/increasing-spaces-between-vertices-for-r-igraph
layout.by.attr <- function(graph, wc, cluster.strength=1,layout=layout.auto) {  
  g <- graph.edgelist(get.edgelist(graph)) # create a lightweight copy of graph w/o the attributes.
  E(g)$weight <- 1

  attr <- cbind(id=1:vcount(g), val=wc)
  g <- g + vertices(unique(attr[,2])) + igraph::edges(unlist(t(attr)), weight=cluster.strength)

  l <- layout(g, weights=E(g)$weight)[1:vcount(graph),]
  return(l)
}

png(filename="instacart_network.png")
set.seed(9) # set seed because orientation of network changes each time
plot(simplify(igraph_adj_network),
     vertex.size = 5,
     vertex.label = ifelse(degree(igraph_adj_network) > 30, V(igraph_adj_network)$label, NA),
     vertex.label.cex = 1,
     vertex.label.font = 2,
     vertex.label.dist = 2,
     vertex.label.color = "black",
     vertex.color = "green4",
     edge.color = "gray81",
     layout=layout.by.attr(igraph_adj_network, wc=1))
title("Network for 5 Instacart orders",
      cex.main = 2)
dev.off()
```



